Macro Impacts

### 2016-08-19 Macro Impacts Revised.tex
Question answered
By wiki on Hankel transform -- the inverse transform has a factor of
(2Pi)^-2

\pagebreak
\section{Thermal Source}
As relayed in an email to Josh, observe that the thermal source is
\begin{align}
T = \f{v_0^2}{c_p}\theta(r_X-r)\,,
\end{align}
where $c_p$ is the specific heat capacity of the material. The equation governing the evolution of the temperature field in 2 dimensions is, with no driving term,
\begin{align}
\partial_t T - \alpha\nabla^2 T = 0
\end{align}
where $\alpha$ is known as the thermal diffusivity, and $\alpha=\f{k}{c_p\rho}$. In cylindrical coordinates and for $T = T(r)$, we have
\begin{align}
\partial_t T - \alpha\nabla_r T = 0
\end{align}
We now Fourier transform in $r$ and $\phi$, yielding
\begin{align}
\partial_t\tilde T =-\alpha k_r^2 \tilde T\,,
\end{align}
where
\begin{align}
\tilde T = 2\pi \int_0^\infty r \diff r  T J_0(k_r r)\,.
\end{align}
Now, we integrate
\begin{align}
\tilde T = C_1e^{-\alpha k_r^2 t}
\end{align}
Now, to determine $C_1$, we set $T = \f{v_0^2}{c_p}\theta(r_X-r)$ at $t = 0$. Observe, this initial condition corresponds to the transformed statement
\begin{align}
\tilde T(0) = \f{r_X}{k_r}J_1(k_r r_X)\,.
\end{align}
This fixes $C_1$. We now inverse transform and obtain
\begin{align}
T &= \int_0^\infty k_r\diff k_r e^{-\alpha k_r^2 t}\f{r_X}{k_r}J_1(k_r r_X) J_0(k_r r)\,,\\
&=r_X\int_0^\infty \diff k_r e^{-\alpha k_r^2 t} J_1(k_r r_X) J_0(k_r r)\,.
\end{align}
We'll have to do this integral numerically. However, we are interested in solving for $r(t)$ such that $T(r(t),t) = T_{\te{melt}}$. We then want to find $\dot r$. Denote this special $r(t)$ as $r_\te{melt}$. Then
\begin{align}
\f{T_\te{melt}}{r_X} = \int_0^\infty \diff k_r e^{-\alpha k_r^2 t} J_1(k_r r_X) J_0(k_r r_\te{melt})
\end{align}
and
\begin{align}
0&=\int_0^\infty \diff k_r \ps{e^{-\alpha k_r^2 t} J_1(k_r r_X) \p{-k_r\dot r_\te{melt}}J_1(k_r r_\te{melt}) +  \p{-\alpha k_r^2}e^{-\alpha k_r^2 t} J_1(k_r r_X)J_0(k_r r_\te{melt})}\,,\\
0&=\int_0^\infty k_r \diff k_r e^{-\alpha k_r^2 t} J_1(k_r r_X) \ps{\p{\dot r_\te{melt}}J_1(k_r r_\te{melt}) +  \p{\alpha k_r}J_0(k_r r_\te{melt})}\,.
\end{align}
\pagebreak\\
Eh... None of this is producing useful results, i.e. those we can use to solve for $r_\te{melt}$. Let's consider a delta source i.e.
\begin{align}
T = \f{v_0^2}{c_p}r_X\delta(r)\,.
\end{align}
Then
\begin{align}
C_1 =2\pi \f{v_0^2}{c_p}r_x
\end{align}
Thus,
\begin{align}
T &= \int_0^\infty k_r\diff k_r e^{-\alpha k_r^2 t}J_0(k_r r)2\pi \f{v_0^2}{c_p}r_x\,,\\
&=\f{r_X v_0^2}{2 c_p t \alpha} e^{-\f{r^2}{4t\alpha}}\,.
\end{align}
(David, what happened to the factor of $2\pi$?)
Setting $T = T_\te{melt}$, we have
\begin{align}
r_\te{melt} = \sqrt{4t\alpha\ln\ps{\f{r_X v_0^2}{2 c_p t \alpha T_\te{melt}}}}\,.
\end{align}
where we've taken the positive branch, since we only care about the melt radius increasing. Then
\begin{align}
\dot r_\te{melt} &=\sqrt{\f{\alpha}{t}}\f{\ln\p{\f{r_Xv_0^2}{2 c_p t\alpha T}} - 1}{\sqrt{\ln\p{\f{r_Xv_0^2}{2 c_p t\alpha T}}}}\,.
\end{align}
For our parameters,  it is a good approximation to take
\begin{align}
\dot r_\te{melt} &=\sqrt{\f{\alpha}{t}\ln\p{\f{r_Xv_0^2}{2 t c_p\alpha T}}}\,.
\end{align}
Thus, the time at which the velocity of the melt front is that of the speed of sound is
\begin{align}
t_{\te{fast}} = \f{\alpha}{v_p^2}\ln\p{\f{r_Xv_0^2}{2 c_p\alpha T_\te{melt}}}
\end{align}
at a radius, which is a good approximation in our case, of
\begin{align}
r_\te{fast melt} =  2\f{\alpha}{v_p}\ln\ps{\f{r_X v_0^2}{2 c_p \alpha T_\te{melt}}}
\end{align}
which is on the order of micrometers to tenths of micrometers for our case.

@@

Consider the initial condition of a delta source along the z-axis in cylindrical coordinates:
\begin{align}
T(r,0) = \f{\abs{\frac{dE}{dx}}}{\rho \sigma_x c_p}\delta(r),
\end{align}
Then, taking the Fourier transform of this initial condition over r and $\phi$ fixes $C_1$ as 
\begin{align}
C_1 =  2\pi \f{\abs{\frac{dE}{dx}}}{\rho \sigma_x c_p}\int_{0}^{\infty}\delta(r)e^{2\pi i k r} dr = 2\pi\f{\abs{\frac{dE}{dx}}}{\rho \sigma_x c_p}\,,
\end{align}
where the factor of $2 \pi$ comes from the $\phi$ integral.
$T$ in real space is given by inverting the Fourier transform:
\begin{align}
T(r,t) &= \int_0^\infty k_r \diff{k_r}\tilde{T} = 4\pi^2\f{\abs{\frac{dE}{dx}}}{\rho \sigma_x c_p}\int_0^\infty k_r\diff k_r e^{-\alpha k_r^2 t}J_0(k_r r)\,,\\
&=\f{\abs{\frac{dE}{dx}}}{2 \rho \sigma_x c_p \alpha t}\ e^{-\f{r^2}{4t\alpha}}\,.
\end{align}
(I'm not sure about the last equality given that I'm still a bit unclear about the factors of $2\pi$ for the inverse transform as noted above after eq. (300) It would seem there are factors from $C_1$ and $\tilde{T}$.  However, the rest of what follows holds, up to some factors of $2\pi$ that hopefully we can clarify).
Setting $T = T_\te{melt}$, we have
\begin{align}
r_\te{melt} = \sqrt{4t\alpha\ln\ps{\f{\abs{\frac{dE}{dx}}}{2 \rho \sigma_x c_p \alpha T_\te{melt} t}}}\,.
\end{align}
where we've taken the positive branch of the square root, since we only care about the melt radius increasing. Then taking the derivative w.r.t to time gives
\begin{align}
\dot r_\te{melt} &=\sqrt{\f{\alpha}{t}}\f{\ln\p{\f{\abs{\frac{dE}{dx}}}{2 \rho \sigma_x c_p \alpha T_\te{melt} t}} - 1}{\sqrt{\ln\p{\f{\abs{\frac{dE}{dx}}}{2 \rho \sigma_x c_p \alpha T_\te{melt} t}}}}\,.
\end{align}
Since, for the parameters at hand, $\ln\p{\f{\abs{\frac{dE}{dx}}}{2 \rho \sigma_x c_p \alpha T_\te{melt} t}}$ is at least of order 10 $>>$ 1, it is a reasonable approximation to take
\begin{align}
\dot r_\te{melt} &=\sqrt{\f{\alpha}{t}\ln\p{\f{\abs{\frac{dE}{dx}}}{2 \rho \sigma_x c_p \alpha T_\te{melt} t}}}\,.
\end{align}
Thus, the time at which the velocity of the melt front is that of the speed of sound, $v_p$, is given by using the Lambert W-function (or "ProductLog" in Mathamatica) as
\begin{align}
t_{\te{fast}} = \f{\alpha}{v_p^2}W\p{{\f{\abs{\frac{dE}{dx}v_p^2}}{2 \rho \sigma_x c_p \alpha^2 T_\te{melt}}}}
\end{align}
Plugging in $t_{\te{fast}}$ to $r_\te{melt}$, which is a solid approximation for the given case, gives
\begin{align}
r_\te{fast melt} =  \frac{2\alpha}{v_p}\sqrt{W\p{{\f{\abs{\frac{dE}{dx}v_p^2}}{2 \rho \sigma_x c_p \alpha^2 T_\te{melt}}}}\ln\ps{\p{\frac{\abs{\frac{dE}{dx}v_p^2}}{2 \rho \sigma_x c_p \alpha^2 T_\te{melt}}}\frac{1}{W\p{{\f{\abs{\frac{dE}{dx}v_p^2}}{2 \rho \sigma_x c_p \alpha^2 T_\te{melt}}}}
}}}\,.
\end{align}
Notice that this has the form of $\sqrt{W(x)\ln\p{{\frac{x}{W(x)}}}}$, which gives $\sqrt{W(x)^2}$ by the identity $\ln\p{{\frac{x}{W(x)}}} = W(x)$ which holds when $x \geq -1/e$ which is true for the parameters at hand.  Thus, we have,
\begin{align}
r_\te{fast melt} =\frac{2\alpha}{v_p}W\p{{\f{\abs{\frac{dE}{dx}v_p^2}}{2 \rho \sigma_x c_p \alpha^2 T_\te{melt}}}}.
\end{align}
which is on the order of micrometers to tenths of micrometers for our case.

>>>>

\pagebreak
\section{Thermal Source}
As relayed in an email to Josh, observe that the thermal source is
\begin{align}
T = \f{v_0^2}{c_p}\theta(r_X-r)\,,
\end{align}
where $c_p$ is the specific heat capacity of the material. The equation governing the evolution of the temperature field in 2 dimensions is, with no driving term,
\begin{align}
\partial_t T - \alpha\nabla^2 T = 0
\end{align}
where $\alpha$ is known as the thermal diffusivity, and $\alpha=\f{k}{c_p\rho}$. In cylindrical coordinates and for $T = T(r)$, we have
\begin{align}
\partial_t T - \alpha\nabla_r T = 0
\end{align}
We now Fourier transform in $r$ and $\phi$, yielding
\begin{align}
\partial_t\tilde T =-\alpha k_r^2 \tilde T\,,
\end{align}
where
\begin{align}
\tilde T = 2\pi \int_0^\infty r \diff r  T J_0(k_r r)\,.
\end{align}
Now, we integrate
\begin{align}
\tilde T = C_1e^{-\alpha k_r^2 t}
\end{align}
Now, to determine $C_1$, we set $T = \f{v_0^2}{c_p}\theta(r_X-r)$ at $t = 0$. Observe, this initial condition corresponds to the transformed statement
\begin{align}
\tilde T(0) = \f{r_X}{k_r}J_1(k_r r_X)\,.
\end{align}
This fixes $C_1$. We now inverse transform and obtain
\begin{align}
T &= \int_0^\infty k_r\diff k_r e^{-\alpha k_r^2 t}\f{r_X}{k_r}J_1(k_r r_X) J_0(k_r r)\,,\\
&=r_X\int_0^\infty \diff k_r e^{-\alpha k_r^2 t} J_1(k_r r_X) J_0(k_r r)\,.
\end{align}
We'll have to do this integral numerically. However, we are interested in solving for $r(t)$ such that $T(r(t),t) = T_{\te{melt}}$. We then want to find $\dot r$. Denote this special $r(t)$ as $r_\te{melt}$. Then
\begin{align}
\f{T_\te{melt}}{r_X} = \int_0^\infty \diff k_r e^{-\alpha k_r^2 t} J_1(k_r r_X) J_0(k_r r_\te{melt})
\end{align}
and
\begin{align}
0&=\int_0^\infty \diff k_r \ps{e^{-\alpha k_r^2 t} J_1(k_r r_X) \p{-k_r\dot r_\te{melt}}J_1(k_r r_\te{melt}) +  \p{-\alpha k_r^2}e^{-\alpha k_r^2 t} J_1(k_r r_X)J_0(k_r r_\te{melt})}\,,\\
0&=\int_0^\infty k_r \diff k_r e^{-\alpha k_r^2 t} J_1(k_r r_X) \ps{\p{\dot r_\te{melt}}J_1(k_r r_\te{melt}) +  \p{\alpha k_r}J_0(k_r r_\te{melt})}\,.
\end{align}
\pagebreak\\
Eh... None of this is producing useful results, i.e. those we can use to solve for $r_\te{melt}$. Let's consider a delta source i.e.
\begin{align}
T = \f{v_0^2}{c_p}r_X\delta(r)\,.
\end{align}
Then
\begin{align}
C_1 =2\pi \f{v_0^2}{c_p}r_x
\end{align}
Thus,
\begin{align}
T &= \int_0^\infty k_r\diff k_r e^{-\alpha k_r^2 t}J_0(k_r r)2\pi \f{v_0^2}{c_p}r_x\,,\\
&=\f{r_X v_0^2}{2 c_p t \alpha} e^{-\f{r^2}{4t\alpha}}\,.
\end{align}
Setting $T = T_\te{melt}$, we have
\begin{align}
r_\te{melt} = \sqrt{4t\alpha\ln\ps{\f{r_X v_0^2}{2 c_p t \alpha T_\te{melt}}}}\,.
\end{align}
where we've taken the positive branch, since we only care about the melt radius increasing. Then
\begin{align}
\dot r_\te{melt} &=\sqrt{\f{\alpha}{t}}\f{\ln\p{\f{r_Xv_0^2}{2 c_p t\alpha T}} - 1}{\sqrt{\ln\p{\f{r_Xv_0^2}{2 c_p t\alpha T}}}}\,.
\end{align}
For our parameters,  it is a good approximation to take
\begin{align}
\dot r_\te{melt} &=\sqrt{\f{\alpha}{t}\ln\p{\f{r_Xv_0^2}{2 t c_p\alpha T}}}\,.
\end{align}
Thus, the time at which the velocity of the melt front is that of the speed of sound is
\begin{align}
t_{\te{fast}} = \f{\alpha}{v_p^2}\ln\p{\f{r_Xv_0^2}{2 c_p\alpha T_\te{melt}}}
\end{align}
at a radius, which is a good approximation in our case, of
\begin{align}
r_\te{fast melt} =  2\f{\alpha}{v_p}\ln\ps{\f{r_X v_0^2}{2 c_p \alpha T_\te{melt}}}
\end{align}
which is on the order of micrometers to tenths of micrometers for our case.

@@

Consider the initial condition of a delta source along the z-axis in cylindrical coordinates:
\begin{align}
T(r,0) = \f{\abs{\frac{dE}{dx}}}{\rho \sigma_x c_p}\delta(r),
\end{align}
Then, taking the Fourier transform of this initial condition over r and $\phi$ fixes $C_1$ as 
\begin{align}
C_1 =  2\pi \f{\abs{\frac{dE}{dx}}}{\rho \sigma_x c_p}\int_{0}^{\infty}\delta(r)e^{2\pi i k r} dr = 2\pi\f{\abs{\frac{dE}{dx}}}{\rho \sigma_x c_p}\,,
\end{align}
where the factor of $2 \pi$ comes from the $\phi$ integral.
$T$ in real space is given by inverting the Fourier transform:
\begin{align}
T(r,t) &=\frac{1}{\p{2\pi}^2} \int_0^\infty k_r \diff{k_r}\tilde{T} = \frac{4\pi^2}{4\pi^2}\f{\abs{\frac{dE}{dx}}}{\rho \sigma_x c_p}\int_0^\infty k_r\diff k_r e^{-\alpha k_r^2 t}J_0(k_r r)\,,\\
&=\f{\abs{\frac{dE}{dx}}}{2 \rho \sigma_x c_p \alpha t}\ e^{-\f{r^2}{4t\alpha}}\,.
\end{align}
Setting $T = T_\te{melt}$, we have
\begin{align}
r_\te{melt} = \sqrt{4t\alpha\ln\ps{\f{\abs{\frac{dE}{dx}}}{2 \rho \sigma_x c_p \alpha T_\te{melt} t}}}\,.
\end{align}
where we've taken the positive branch of the square root, since we only care about the melt radius increasing. Then taking the derivative w.r.t to time gives
\begin{align}
\dot r_\te{melt} &=\sqrt{\f{\alpha}{t}}\f{\ln\p{\f{\abs{\frac{dE}{dx}}}{2 \rho \sigma_x c_p \alpha T_\te{melt} t}} - 1}{\sqrt{\ln\p{\f{\abs{\frac{dE}{dx}}}{2 \rho \sigma_x c_p \alpha T_\te{melt} t}}}}\,.
\end{align}
Since, for the parameters at hand, $\ln\p{\f{\abs{\frac{dE}{dx}}}{2 \rho \sigma_x c_p \alpha T_\te{melt} t}}$ is at least of order 10 $>>$ 1, it is a reasonable approximation to take
\begin{align}
\dot r_\te{melt} &=\sqrt{\f{\alpha}{t}\ln\p{\f{\abs{\frac{dE}{dx}}}{2 \rho \sigma_x c_p \alpha T_\te{melt} t}}}\,.
\end{align}
Thus, the time at which the velocity of the melt front is that of the speed of sound, $v_p$, is given by using the Lambert W-function (or "ProductLog" in Mathamatica) as
\begin{align}
t_{\te{fast}} = \f{\alpha}{v_p^2}W\p{{\f{\abs{\frac{dE}{dx}v_p^2}}{2 \rho \sigma_x c_p \alpha^2 T_\te{melt}}}}
\end{align}
Plugging in $t_{\te{fast}}$ to $r_\te{melt}$, which is a solid approximation for the given case, gives
\begin{align}
r_\te{fast melt} =  \frac{2\alpha}{v_p}\sqrt{W\p{{\f{\abs{\frac{dE}{dx}v_p^2}}{2 \rho \sigma_x c_p \alpha^2 T_\te{melt}}}}\ln\ps{\p{\frac{\abs{\frac{dE}{dx}v_p^2}}{2 \rho \sigma_x c_p \alpha^2 T_\te{melt}}}\frac{1}{W\p{{\f{\abs{\frac{dE}{dx}v_p^2}}{2 \rho \sigma_x c_p \alpha^2 T_\te{melt}}}}
}}}\,.
\end{align}
Notice that this has the form of $\sqrt{W(x)\ln\p{{\frac{x}{W(x)}}}}$, which gives $\sqrt{W(x)^2}$ by the identity $\ln\p{{\frac{x}{W(x)}}} = W(x)$ which holds when $x \geq -1/e$ which is true for the parameters at hand.  Thus, we have,
\begin{align}
r_\te{fast melt} =\frac{2\alpha}{v_p}W\p{{\f{\abs{\frac{dE}{dx}v_p^2}}{2 \rho \sigma_x c_p \alpha^2 T_\te{melt}}}}.
\end{align}
which is on the order of micrometers to tenths of micrometers for our case.
